Hysteresis in human computation:\\ how one task affects another

### 2014-11-25 docs/publications/2014.09_SCIENCE/man.tex
significant portion of supplementary material done

\usepackage{graphicx}
\usepackage{amsthm}
\newtheorem{mydef}{Definition}
\usepackage{dcolumn}
\usepackage{multirow}
\usepackage{booktabs}
\newcolumntype{d}{D{.}{.}{4.0}}
\newcolumntype{s}{D{.}{.}{1.4}}

@@

\section*{Introduction}
Microtask crowdsourcing platforms like Amazon Mechanical Turk (MTurk) make it 
possible for requesters to submit batches of tasks to a large pool of 
workers, who do the tasks for fun, a sense of purpose, and remuneration 
\cite{kazai2013analysis,Antin20122925}.  
Typical tasks include tagging and categorizing images 
\cite{6116320,Zhai2012357}, transcribing voice recordings 
\cite{chandler2013breaking,paolacci2010running}
or handwritten notes \cite{Berinsky2012351,Finnerty2013}, and judging the 
relevancy of search results 
\cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}.
Originally used to distribute clerical work, these platforms 
increasingly serve as a means to engage experimental participants in a 
research 
setting \cite{paolacci2010running,Berinsky2012351,snow2008cheap,alonso2009can}.
This trend has seeded considerable interest in characterizing the parameters
of task design, and validating the microtask platform methodology (\#).

Computer scientists view microtask platforms in the context of general
computation.  In this regard, the platforms enable a tight integration of 
human and machine computation.  Researchers coined the term HPU 
(Human co-Processing Unit) in analogy to CPUs and GPUs \cite{5543192}.  
Ongoing research pursues the establishment of an HPU instruction set
including basic operators like \texttt{vote}, \texttt{subdivide}, \texttt{map},
\texttt{reduce}, \texttt{sentiment}, and \texttt{image\_tag} (\#).  Libraries 
are being developed to give the programmer more seamless access to HPU 
resources 
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.  
As with other architectures, the HPU architecture makes entirely new classes 
of problems tractable for the programmer (\#).

But the ability to program with high-level human instructions comes with a 
cost: HPU output is inherently noisy.  Part of this noise can be explained
in terms of worker demography (\#).  But systematic cognitive biases also 
play a role (\#).   
So far, investigation has explored how worker performance is influenced by
such design factors as the level of 
pay \cite{kazai2013analysis}, training \cite{le2010ensuring}, pre-screening of 
workers \cite{paolacci2010running}, and user-interface design 
\cite{Finnerty2013}.  Researchers have also investigated \textit{framing}, 
by testing the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or using alternate problem descriptions
\cite{thibodeau2013natural}.  To our knowledge, no study has investigated 
the influences that consecutive microtasks exert on one another.

@@

Here we describe a series of experiments, using the Amazon Mechanical
Turk (MTurk) microtask platform, in which we studied intertask effects
for image-labeling tasks.  Image-labeling tasks are among the most 
common kinds of microtasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, 
and it seems likely that this will remain the case in the long term because
they provide a source of ground truth in computer vision research 
\cite{5543192}.  
During the experiments, workers were presented with a series of images (one at
a time), and had to provide five descriptive words (\textit{labels}) for each.
We varied first five images (hereafter the \textit{initial tasks})
and observed the effect this had on the workers' labels for the last five
(\textit{test tasks}).

The initial tasks were chosen to embed a certain \textit{prime concept}.
In our first experiment, our prime concepts were \textit{food} and (non-food) 
\textit{objects}.  In our second experiment, the prime concepts were 
\textit{food} and \textit{culture}.  The test tasks for the experiments used 
images that combined both prime concepts used in the initial tasks. Workers 
participated in only one treatment of one experiment, and were randomly 
assigned initial tasks based on one of the 
prime concepts.  From the perspective of the worker, we made no distinction 
between the initial tasks and test tasks.

As a point of comparison, we included treatments in which we induced
\textit{framing}.
In the framing treatments, before workers begin the tasks, we told them that 
the purpose of study was to understand the visual perception of 
``food and ingredients'' in one case, or the 
perception of ``objects and tools'' in another.  We varied the language of 
the framing text (hereafter simply the \textit{frame}), and experimented with 
requiring the worker to echo the frame using a combo-box input.

@@

\paragraph{Experimental setup.}
\begin{figure}
	\centering
	\includegraphics[scale=0.7]{figs/images.jpg}
	\caption{
		Examples of images used in labeling tasks. (The italicized treatment 
		names that follow are defined in Table \ref{table:treatments}.) Images 
		from initial tasks: A) \textit{task1:food}, B) 
		\textit{task1:obj}, D) \textit{task2:food}, 
		E) \textit{task2:cult}.  Images from  test tasks: C) \textit{exp1}, 
		F) \textit{exp2}.  The full set of experimental materials is 
		shown in the supplementary material.
	}

@@

Note that food and objects are both relatively well-defined classes of 
physical entities.  To explore a different conceptual axis, in 
\textit{exp2} we adopted the prime concepts \textit{food} and \textit{culture}.
Compared to food and objects, culture is more abstract, and, rather than 
comprising a class of physical entities, it encapsulates
sets of beliefs, values, and practices.  Examples of images from the initial 
tasks in \textit{task2} are shown in Fig.~\ref{fig:task}, and the full 
collection of 
experimental materials is presented in the supplementary material.  We 
ensured that all workers only participated in one treatment of one experiment.

\setlength{\tabcolsep}{3pt}
\begin{table}[t]
\centering
\begin{tabular}{ c c c c c }
		\hline \noalign{\smallskip}
		\multicolumn{3}{c}{\textbf{Treatment name}} & \parbox[c]{1.6cm}{\centering \textbf{Frame}} & \parbox[c]{1.3cm}{\centering \textbf{Initial\\ tasks}}	\\

@@

\subsection*{Assessing task similarity}
\begin{table}
\centering
\begin{tabular}{ l  s s s s}

\toprule    
Image set   
& \multicolumn{1}{c}{Ambig.} 
& \multicolumn{1}{c}{Cultural} 
& \multicolumn{1}{c}{Ingr.}
& \multicolumn{1}{c}{Test} \\

\midrule

Ambiguous  & 1 & 0.0418 & 0.142 & 0.167 \\

Cultural  & 0.0418  & 1 & 0.0347 & 0.0561 \\

Ingredients  & 0.142  & 0.0347 & 1 & 0.110 \\

Test & 0.167  & 0.0561 & 0.110 & 1
\\
\bottomrule

\end{tabular}
\caption{\footnotesize{
Pairwise similarities of each image set based on the labels attributed to them (see \textbf{Eq. 4}).
}}
\label{table:2}
\end{table}

One weakness of the above analysis is the reliance on the informal notion
of the similarity of one set of images to another. The 
characterization of image content is a deeply complex issue that has been 
approached by many disciplines \cite{panofsky1939studies,shatford1986analyzing,Tversky1977327,Jaimes20002}. \textit{Perceptual similarity} is a 
particularly recalcitrant concept.   But to formalize our claim, we seek a 
measure of similarity between two sets of images, 
\textit{with respect to the labeling task}.  This can be operationalized
in a straightforward way, by using the similarity of the labels given to images
as a proxy for the similarity of the images.  Thus, to measure the 
similarity between two sets of images, X and Y , we computed the Jacquard 
index between the sets of labels attributed to them:
$$
	\mathrm{Sim}(X,Y) = \frac{L(X) \cap L(Y)}{L(X) \cup L(Y)}
$$
where $L(X)$ denotes the set of labels attributed to images in $X$.

The pairwise similarities of the image sets are presented in Table 2. This 
shows that the workers that yielded higher specificity labels were from 
treatments whose initial tasks were more similar to the test tasks (c.f.
\textbf{Fig 3}).

Such a phenomenon would be consistent with the psychological mechanism known as
\textit{negative priming}.

>>>>

\usepackage{framed, color}
\usepackage{soul}
\definecolor{blu}{rgb}{0,0,1}
\newcommand{\td}[1]{{\color{blu}\hl{TODO: #1}}}
\usepackage{graphicx}
\usepackage{amsthm}
\newtheorem{mydef}{Definition}
\usepackage{dcolumn}
\usepackage{multirow}
\usepackage{booktabs}
\newcolumntype{d}{D{.}{.}{4.0}}
\newcolumntype{s}{D{.}{.}{1.4}}

@@

\section*{Introduction}
Microtask crowdsourcing platforms like Amazon Mechanical Turk (MTurk) make it 
possible for requesters to submit batches of tasks to a large pool of 
workers, who do the tasks for fun, a sense of purpose, and remuneration 
\cite{kazai2013analysis,Antin20122925}.  
Typical tasks include tagging and categorizing images 
\cite{6116320,Zhai2012357}, transcribing voice recordings 
\cite{chandler2013breaking,paolacci2010running}
or handwritten notes \cite{Berinsky2012351,Finnerty2013}, and judging the 
relevancy of search results 
\cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}.
Originally used to distribute clerical work (\#), these platforms 
increasingly serve as a means to engage experimental participants in a 
research 
setting \cite{paolacci2010running,Berinsky2012351,snow2008cheap,alonso2009can}.
This trend has seeded considerable interest in characterizing the parameters
of task design, and validating the microtask platform methodology (\#).

Computer scientists view microtask platforms in the context of general
computation(\#).  In this regard, the platforms enable a tight integration of 
human and machine computation(\#).  Researchers coined the term HPU 
(Human co-Processing Unit) in analogy to CPUs and GPUs \cite{5543192}.  
Ongoing research pursues the establishment of an HPU instruction set
including basic operators like \texttt{vote}, \texttt{subdivide}, \texttt{map},
\texttt{reduce}, \texttt{sentiment}, and \texttt{image\_tag} (\#).  Libraries 
are being developed to give the programmer more seamless access to HPU 
resources 
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.  
As with other architectures, the HPU architecture makes entirely new classes 
of problems tractable for the programmer (\#).

But the ability to program with high-level human instructions comes with a 
cost: HPU output is inherently noisy (\#).  Part of this noise can be explained
in terms of worker demography (\#).  But systematic cognitive biases also 
play a role (\#).   
So far, investigation has explored how worker performance is influenced by
such design factors as the level of 
pay \cite{kazai2013analysis}, training \cite{le2010ensuring}, pre-screening of 
workers \cite{paolacci2010running}, and user-interface design 
\cite{Finnerty2013}.  Researchers have also investigated \textit{framing}, 
by testing the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or using alternate problem descriptions
\cite{thibodeau2013natural}.  To our knowledge, no study has investigated 
the influences that consecutive microtasks exert on one another.

@@

Here we describe a series of experiments, using the Amazon Mechanical
Turk (MTurk) microtask platform (\#), in which we studied intertask effects
for image-labeling tasks.  Image-labeling tasks are among the most 
common kinds of microtasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, 
and it seems likely that this will remain the case in the long term because
they provide a source of ground truth in computer vision research 
\cite{5543192}.  
During the experiments, workers were presented with a series of images (one at
a time), and had to provide five descriptive words (\textit{labels}) for each.
We varied first five images (hereafter the \textit{initial tasks})
and observed the effect this had on the workers' labels for the last five
(\textit{test tasks}).

The initial tasks were chosen to embed a certain \textit{prime concept}.
Our first experiment compares the prime concepts \textit{food} and (non-food) 
\textit{objects}, while our second experiment compares \textit{food} and 
\textit{culture}.  For both experiments, The test tasks used images that
combined both prime concepts used for the initial tasks \td{reword}.  Workers participated 
in only one treatment of one experiment, and were randomly assigned initial 
tasks based on one of the prime concepts.  From the perspective of the worker, 
we made no distinction between the initial tasks and test tasks.

As a point of comparison, we included treatments in which we induced
\textit{framing}. \td{what is framing?}
In the framing treatments, before workers begin the tasks, we told them that 
the purpose of study was to understand the visual perception of 
``food and ingredients'' in one case, or the 
perception of ``objects and tools'' in another.  We varied the language of 
the framing text (hereafter simply the \textit{frame}), and for some treatments
we required the worker to echo the frame using a combo-box input.

@@

\paragraph{Experimental setup.}
\begin{figure}
	\centering
	\includegraphics[scale=0.7]{figs/images.pdf}
	\caption{
		Examples of images used in labeling tasks. (The italicized treatment 
		names that follow are defined in Table \ref{table:treatments}.) Images 
		from initial tasks: A) \textit{task1:food}, B) 
		\textit{task1:obj}, D) \textit{task2:food}, 
		E) \textit{task2:cult}.  Images from  test tasks: C) \textit{exp1}, 
		F) \textit{exp2}.  The full set of experimental materials is 
		shown in the supplementary material (S1).
	}

@@

Note that food and objects are both relatively well-defined classes of 
physical entities.  To explore a different conceptual axis, in 
\textit{exp2} we adopted the prime concepts \textit{food} and \textit{culture}.
Compared to food and objects, culture is more abstract, and, rather than 
comprising a class of physical entities, it encapsulates
sets of beliefs, values, and practices.  Examples of images from the initial 
tasks in \textit{task2} are shown in Fig.~\ref{fig:task}, and the full 
collection of 
experimental materials is presented in the supplementary material (S1).  We 
ensured that all workers only participated in one treatment of one experiment.

\setlength{\tabcolsep}{2pt}
\begin{table}[t]
\centering
\begin{tabular}{ c c c c c }
		\hline \noalign{\smallskip}
		\multicolumn{3}{c}{\textbf{Treatment name}} & \parbox[c]{1.6cm}{\centering \textbf{Frame}} & \parbox[c]{1.3cm}{\centering \textbf{Initial\\ tasks}}	\\

@@

\subsection*{S1: Experimental materials}
The tasks for both experiments were presented as a series of slides.  In
both cases, the first slide consisted of a brief set of instructions, followed
by the frame if shown (only some treatments involved a frame), see 
Fig. S\ref{fig:hit_preamble}.  Following the instructions and prime, a set
of initial tasks was shown (except for the treatments in \textit{frame2}
and \textit{echo}) followed by (in all cases) a set of test tasks.  In
\textit{exp1}, two kinds of frame were used, one simple frame, and one where
the worker was required to echo the frame using a combo-box input.  The
instructions, both kinds of frame, and an example task slide, are shown 
in Fig. S1.
\begin{figure}
	\includegraphics[scale=0.8]{figs/tasks.pdf}
	\label{fig:hit_preamble}
	\caption{Examples of A) instructions; B) a frame, as shown in 
		\textit{frame1:obj}, and similar to those shown in 
		\textit{frame1:food} and \textit{frame2:food}, and 
		\textit{frame2:cult};  
		C) an echoed frame, as used in \textit{echo:obj}, and similar to that
		used in \textit{echo:food}; and D) an 
		example of an image labeling task.
	}
\end{figure}

All of the initial tasks had the format shown in Fig. S1D.  The images used
in the initial tasks for \textit{task1:food} and \textit{task1:obj} are
shown in Fig. S2 and S3, and the images used in the test tasks for 
\textit{exp1} are shown in Fig. S4.  The images used in the initial tasks
for \textit{task2:food} and \textit{task2:cult} are shown in Fig. S5 and
Fig. S6 respectively, and the images used in the test tasks for \textit{exp2} 
are shown in Fig. S7.

\begin{figure}
	\includegraphics{figs/task1-food.pdf}
	\label{fig:task1:food}
	\caption{
		Figure S2: Images used in the initial tasks for 
		\textit{task1:food}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task1-obj.pdf}
	\label{fig:task1:obj}
	\caption{
		Figure S3: Images used in the initial tasks for 
		\textit{task1:obj}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task1-test.pdf}
	\label{fig:task1:test}
	\caption{
		Figure S2: Images used in the test tasks for 
		\textit{exp1}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task2-food.pdf}
	\label{fig:task2:food}
	\caption{
		Figure S2: Images used in the initial tasks for 
		\textit{task2:food}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task2-cult.pdf}
	\label{fig:task2:cult}
	\caption{
		Figure S2: Images used in the initial tasks for 
		\textit{task2:cult}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task2-test.pdf}
	\label{fig:task2:test}
	\caption{
		Figure S2: Images used in the test tasks for 
		\textit{exp2}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\subsection*{S2: Access to raw data and code}
The raw data and source code needed to reproduce the analyses 
shown here and in the main text can be downloaded from 
http://networkdynamics.org/publications/public/priming2014.09.

\subsection*{S3: Additional results not shown in the main text}
\begin{itemize}
	\item{
		the top 5 most suppressed and most activated terms between treatment
		pairs.
	}
\end{itemize}

\subsection*{S4: Data preprocessing}
	\paragraph{Splitting, Lemmatization, removal of stop-words, and 
		addition of position tags.}

	Before performing any analysis on the labels that workers provided, we
	performed a series of preprocessing steps.  
	\td{does this reflect the actual order in which the steps were done?}
	First, words wore lematized 
	using the
	\td{whatzit} lemmatizer.  This lemmatizer is very mild, tending only to 
	change plural words to their singular form.  Common
	words such as ``the'', ``to'', or ``with'', were found using the
	the \td{whatzit} stop-word list and removed.  Then, misspelled words were
	automatically corrected using a spelling correction algorithm described 
	below.  Finally, labels that contained
	multiple words (separated by spaces or punctuation) were split, with
	punctuation removed, and the separated words were treated as distinct 
	features in subsequent analysis.

	For the purpose of training and testing a naive Bayes classifier, we 
	performed an additional preprocessing step.  We tagged words with the
	position in which they had been entered (i.e. which of the five text 
	inputs in the task interface) as well as the test tasks in which the word 
	had been provided.
	So, if the word ``wine'' was entered into the second text-input for 
	the third test-task, after preprocessing, the feature ``3\_2\_wine'' would
	appear.  Prepending the task number onto words was simply a means to 
	retain correct attribution, which itself was essential since providing a
	word during one task is not equivalent to providing the same word during 
	another task.

	\paragraph{Spelling correction.}  
	Spelling correction was performed using an algorithm that first detected
	if a word was likely to be misspelled, then generated a set of candidate 
	corrections, and chose the best candidate based on a scoring mechanism.

	A word was considered misspelled if it was not contained in the 
	\textit{legal set}, which was formed by the union of
	the wordnet corpus, the stop-word list, and a set of words seen while 
	crawling the world food section of the allrecipes.com website.  We
	describe the crawling of the allrecipes.com website in a section below.

	To correct misspellings, we first produced all possible modified forms 
	that could be obtained applying one or two edits.  An edit consisted of 
	adding or removing a letter, changing one letter into another, or 
	swapping the positions of two adjacent letters.  For the purpose of these 
	edits, spaces were treated as any other letter.

	The candidates produced by these edits which were in the legal set were
	then ranked based on a scoring mechanism, and the highest scoring word
	was chosen as the correction.  A candidate $w$'s score, $s_w$, was 
	calculated according to the following formula:
	\begin{equation}
		s_w = (f_w + 1) * p_1 * p_2,
	\end{equation}
	where $f_w$ is the frequency with wich the word occured (correctly 
	spelled)
	within the given task, $p_1$ was a penalty for the first edit, and
	$p_2$ was a penalty for the second edit.  If the word was made using only
	one edit, then $p_2 = 1$.  Any edit that did not involve adding a space
	(i.e. separating a word) incured a penalty of 0.5, while the addition of
	a space incurred a penalty of 0.1.  Word-separation edits were more 
	strongly penalized because it is often possible to split a series of 
	letters into many short two- or three-letter words, which otherwise leads 
	to many erroneous corrections.  We found that setting the penalty for
	word splitting at 0.1 worked well by testing on words taken from
	initial tasks.

	After all of our analyses had been performed, we checked the accuracy of 
	the spelling correction algorithm using three human coders.  The coders
	were shown a set of 
	500 words that were randomly sampled from the labels attirubuted during 
	test tasks (50 for each test task).  Before sampling, the labels from all
	experimental treatments, for the given test task, were pooled together.
	The coders did not know which treatment any given word came from.  
	In addition to the words, coders were shown the spelling correction 
	produced by the algorithm, as well as the image from the test task.
	The coders were asked to identify any misspelled words which had not 
	been corrected, as well as any corrections that appeared to be erroneous,
	by indicating their own correction.

	Based on this exercise, we found that, according to the human coders,
	17\% of the words were misspelled, but that after applying the 
	automated spelling correction, only 3.2\% of words were misspelled.

	\paragraph{Crawling the world food section of allrecipes.com.}
	The website allrecipes.com was accessed on 3-4 November 2014 using 
	automated scripts.  A total of 2642 recipe listings and 15621 recipe
	instruction pages were crawled.  Recipe listings were pages that provided 
	lists of recipes, and contained a title and short description for each.  
	The recipe instruction pages had lists of ingredients and preparation 
	instructions.  All of the words found in recipe titles, short
	descriptions, ingredients lists, and preparation instructions were
	collected, and saved as an auxiliary set to aument the \textit{legal set}
	used in spelling correction.

\subsection*{S5: Identification of food-related terms}
The wordnet corpus is composed of synsets, which are particular senses 
(meanings) and a set of word forms bearing that meaning. A given word form, 
such as ``ring'' has multiple senses (``to ring a bell'', ``a wedding ring''), 
and so can be part of many synsets.  Since the wordnet corpus was designed
to carry semantic information, the synset is the basic organizing element
of wordnet.

We chose the synsets \textit{food.n.01} and \textit{food.n.02} to act as roots
in defining which words should be considered food-related.  Any word form
belonging to a synset which was a hyponym of one of the two root synsets
identified above was considered to be a reference to food.  Here, as in the
main text, when we say hyponym, we mean either a direct hyponym, or a hyponym
of a hyponym, and so on.

This is a stringent notion of a word being a food reference.  It roughly
corresponds to whether or not a given thing is reasonably considered 
consumable.  So, ``orange'' is food, but ``salty'' is not.  Although ``salty''
would usually refer to something edible, it is not itself an edible thing.
On the other hand, ``salt'' would be considered a reference to food under 
this operationalization.

The subset of the wordnet corpus induced by the hyponyms of \textit{food.n.01}
and \textit{food.n.02} provide an extensive set of \td{X} words.  Manual
testing appeared to show that the coverage was very extensive, except in 
some cases references to ethnic foods were missing.  Since, especially in
\textit{exp2} we had images containing many ethnic foods, we decided to 
augment the wordnet corpus with words learned by crawling the world food
section of the allrecipes.com website.

After collecting a set of words using an automated script 
(\td{described above}), we collected the set of words that had been used by
workers and which had been seen during crawling of the allrecipes.com website,
but which were not included in wordnet.  We graphted these extra words into
the wordnet hyponym-hypernym graph manually, which ensured that any references
to ethnic words learned from crawling would be properly classified as 
food references.

\paragraph{Validation of the detection of food references.}
To determine whether our extended version of wordnet provided a good approach
to detecting food-related words, we sampled 500 words from the set of labels
produced by workers and had three human coders manually decide if they were 
food-related or not.  The method for randomly sampling the 500 words was 
described in above in \td{section number}.

The coders were instructed to consider whether a word was a noun signifying
an edible item.  The guiding principle was, for the coder to ask herself,
``can I eat this X'', where X is the word she is coding.  The coders were
shown the used in the task from which the word came, to help resolve ambiguity
in the sense of the word that had been given.  Coders were also shown the 
spelling correction (if any) that the spell-correction algorithm had made,
to help interpret mispelled words \td{we describe the spelling correction 
algorithm in section X}.

We evaluated the food detection algorithm in two ways.  First, we measured
the inter-rater reliability between the three human coders and the algorithm
(treating the algorithm just like any other coder).  Second, adopting the 
majority code given by the human coders, we determined the accuracy of the
food-detection algorithm.  Data sumarizing this validation process are given
in table S1.

\begin{table}
\centering
\setlength{\tabcolsep}{12pt}
\begin{tabular}{ r | c }
\toprule    
\# terms coded & 500 \\
\# food references & 130 \\
\# correct machine codes & 440 \\
inter-rater reliability & 82.4\% \\
human-machine code agreement & 88\% \\
%machine recall & 80 \% \\
%machine precision & 75.4 \% \\
%machine code F1 & 0.78 \\
\bottomrule
\end{tabular}
\caption{\footnotesize{
	\td{remove the automatic numbering}. Table S1. Results for the validation 
	of the automatic detection of food references, by comparison to
	human-generated coding on randomly sampled words.  In calculating the 
	human-machine code agreement, the majority human code was used.
	Inter-rater reliability was calculated as Krippendorff's alpha.
}}
\end{table}

The results show that good inter-rater reliability was achieved during the
validation (82.4\%).  Looking at the agreement between the 
machine codes and the majority human-generated codes, we find they agree
88\% of the time.  This means that the concept used to instruct human coders
provides a good approximation to what the wordnet-based machine codes actually
represent. In other words, the machine codes correspond closely to indicating
which words correspond to nouns that refer to edible things, and which ones
don't, which provides a clear and simple interpretation for the machine
coding of food and non-food words.

\subsection*{S6: Calculation of relative specificity}
In the main text we present results for the relative specificity of labels
produced by two experimental treatments.  In performing this calculation,
we used the wordnet corpus, which has a set of hypernym-hyponym relationships
as described in the main text.  The first step in performing this calculation
was therefore to map the words occuring from (the labels of) a given 
experimental treatment onto the wordnet synsets.  In general, a given word
form can have multiple meanings, and therefore maps onto multiple wordnet 
synsets.

The result of mapping the words from two treatments onto the wordnet 
hypernym-hyponym graph two sets of counts, one set per treatment, 
indicating the number of times a word from a given synset occured among the
labels of the treatment.  Then, for every synset in one treatment, we looked
at the number of synsets in the other treatment that were more generic
(reachable by following hypernym relations) less the number of words that
were more specific (reachable by following hyponym relations).  This quantity
tallied for all words in the original synset gives a (non-normalied) measure
of the degree to which words from the first treatment are more specific
than words from the second treatment, overall.  We then normalized this 
quantity by the total number of possible comparisons between the words of
one treatment to those of the other, where two words are considered comparable
only if one word is the hypernym or hyponym of the other.  Hence, ``statue'' 
and ``bread'' are not comparable, but ``pumpernickel'' and ``bread'' are.

This calculation can be summarized by the following equation:
\begin{equation}
	S(P,Q) = \frac{
		\sum_{w\in P}\sum_{v\in Q} \left(
			\mathbf{1}_{[w>v]} - \mathbf{1}_{[v>w]} \right)
	}{
		\sum_{w\in P}\sum_{v\in Q} \left(
			\mathbf{1}_{[w>v]} + \mathbf{1}_{[v>w]} \right)
	},
\end{equation}
where $P$ and $Q$ are sets of synsets, and $\mathbf{1}_{[w>v]}$ evaluates
to 1 if synset $w$ is more specific than (i.e. is a hyponym of) synset $v$.
This measure counts the excess of cases where synsets from $P$ are more
specific then synsets from $Q$, as a fraction of all comparable synset pairs.

In computing this quantity between two treatments, we first computed it 
using the labels attributed to a particular test task, and then
averaged the result obtained in this way for each of the five test tasks.

\subsection*{S7: Analysis of variance}
\begin{itemize}
	\item{
		classifier accuracy.  Why we used a one-tail CI.  
		Why certain error bars are smaller.
	}
	\item{error in $\delta$ \%food.}
	\item{error in size of food vocabulary}
	\item{error in food specificity}
\end{itemize}

\subsection*{S8: Measuring computational hysteresis}
\begin{itemize}
	\item{principled use of cross-validation without optimization.}
	\item{Adopting Naive Bayes due to the lack of optimization required.}
	\item{
		Usage of a principled set of pre-processing and representation 
		techniques.
	}
	\item{
		Comparison to other choices of representation
	}
	\item{
		Comparison to SVM classifier
	}
	\item{
		Reiterate why we cannot report optimized results.
	}
\end{itemize}

\subsection*{S9: Justifications and statistical properties of classifier-based measurement of computational hysteresis}
\begin{itemize}
	\item{Relationship to L1-Distance}
	\item{Relationship to bias}
	\item{Relationship to Jenson-Shannon divergence}
	\item{Other approaches to measuring L1-distance}
	\item{Comparison to hypothesis testing using $\chi^2$}
\end{itemize}

%\subsection*{Assessing task similarity}
%\begin{table}
%\centering
%\begin{tabular}{ l  s s s s}
%
%\toprule    
%Image set   
%& \multicolumn{1}{c}{Ambig.} 
%& \multicolumn{1}{c}{Cultural} 
%& \multicolumn{1}{c}{Ingr.}
%& \multicolumn{1}{c}{Test} \\
%  
%\midrule
%
%Ambiguous  & 1 & 0.0418 & 0.142 & 0.167 \\
%
%Cultural  & 0.0418  & 1 & 0.0347 & 0.0561 \\
%
%Ingredients  & 0.142  & 0.0347 & 1 & 0.110 \\
%
%Test & 0.167  & 0.0561 & 0.110 & 1
%\\
%\bottomrule
%
%\end{tabular}
%\caption{\footnotesize{
%Pairwise similarities of each image set based on the labels attributed to them (see \textbf{Eq. 4}).
%}}
%\label{table:2}
%\end{table}
%
%One weakness of the above analysis is the reliance on the informal notion
%of the similarity of one set of images to another. The 
%characterization of image content is a deeply complex issue that has been 
%approached by many disciplines \cite{panofsky1939studies,shatford1986analyzing,Tversky1977327,Jaimes20002}. \textit{Perceptual similarity} is a 
%particularly recalcitrant concept.   But to formalize our claim, we seek a 
%measure of similarity between two sets of images, 
%\textit{with respect to the labeling task}.  This can be operationalized
%in a straightforward way, by using the similarity of the labels given to images
%as a proxy for the similarity of the images.  Thus, to measure the 
%similarity between two sets of images, X and Y , we computed the Jacquard 
%index between the sets of labels attributed to them:
%$$
%	\mathrm{Sim}(X,Y) = \frac{L(X) \cap L(Y)}{L(X) \cup L(Y)}
%$$
%where $L(X)$ denotes the set of labels attributed to images in $X$.
%
%The pairwise similarities of the image sets are presented in Table 2. This 
%shows that the workers that yielded higher specificity labels were from 
%treatments whose initial tasks were more similar to the test tasks (c.f.
%\textbf{Fig 3}).
%
%Such a phenomenon would be consistent with the psychological mechanism known as
%\textit{negative priming}.
